It has been quite an enjoyed data mining journey from the beginning up to this point. Haven taken this data mining journing thus far, the following the following recommendation(s) and conclusion are here by made:
\begin{description}
	\item[Recommendation] I recommend that :  
	\begin{itemize}
		\item Data should be collect about how passionat the students are about their programme / course of study.
		
		\item Data should be collected about how easy the students feel their courses.
		
		\item Data should be collect about how much the students love their teacher and how much they feel the teacher loves and cares for their well-being in and outside of the classroom
		
		\item Data about how much the academic works related to students' inherent talents should also be collected and wrangled.
	\end{itemize}

	\item Conclusion
	\begin{itemize}
		\item From the trends and or patterns in the data, student grade or performance does not depend on his or her gender.
		
		\item Computer Science (CS) is best performming department possibly because for more than two decades, Data Science, Machine Learning, Artificial Intelligence have been trending and getting people's attension. This and availability of learning resourse on youtube, stackoverflow and others are some of the reasons the departent of Computer Science is the best performing department.
		
		\item Students whose parents are Bachelor's Degree holder performed best followed by the students whose parents are Ph.D holders.
		
		\item In all grades, students whose parents are high  income earners took the lead.
	\end{itemize}
\end{description}