\pagestyle{fancy}
\section{Data Collection \& Understanding}%\label{sec:background}
The dataset, titled \bfseries{Student Performance \& Behaviour Dataset} is a dataset of 5000 records collected from a private learning provider. It contains 5000 records and 23 features.  \cite{StudentPerformanceBehavior} of different and diverse students' behaviors and environmental factors that go a long way, directly and or indirectly affecting a students academic performance. And it has key student attributes necessary for exploring patterns, correlations and insights related to academic performance. Below is the data dictionary table 

\begin{enumerate}
	\item \textbf{Data Dictionary}
	\begin{enumerate}
		\item Student\_ID: Unique identifier for each student
		\item First\_Name: Student's first name 
		\item Last\_Name: Student's last name 
		\item Email: Contact email (can be anonymized)
		\item Gender: Male, Female and Other
		\item Age: The age of the student
		\item Department: Student's department (e.g., CS, Engineering, Business).
		\item Attendance (\%): Attendance percentage (0-100%).
		\item Midterm\_Score: Midterm exam score (out of 100).
		\item Final\_Score: Final exam score (out of 100).
		\item Assignments\_Avg: Average score of all assignments (out of 100).
		\item Quizzes\_Avg: Average quiz scores (out of 100).
		\item Participation\_Score: Score based on class participation (0-10).
		\item Projects\_Score: Project evaluation score (out of 100).
		\item Total\_Score: Weighted sum of all grades.
		\item Grade: Letter grade (A, B, C, D, F).
		\item Study\_Hours\_per\_Week: Average study hours per week.
		\item Extracurricular\_Activities: Whether the student participates in extracurriculars (Yes/No).
		\item Internet\_Access\_at\_Home: Does the student have access to the internet at home? (Yes/No).
		\item Parent\_Education\_Level: Highest education level of parents (None, High School, Bachelor's, Master's, PhD).
		\item Family\_Income\_Level: Low, Medium, High.
		\item Stress\_Level (1-10): Self-reported stress level (1: Low, 10: High).
		\item Sleep\_Hours\_per\_Night: Average hours of sleep per night.
		
	\end{enumerate} 	
\end{enumerate}



%\subsection{Medicinal Properties}
%\subsection{Preservation methods}
%\subsection{Phytochemical content of Ugu}

%\section{Problem Statement}
%\section{Objective of the Study}
%\section{Research Questions}
%\section{Justification of the study} % (Significance of the study)}
%\section{Scope of the Study}